\large \texttt{insectDisease}: programmatic access to the \textit{Ecological Database of the World's Insect Pathogens}

### 2022-04-18 manuscript/manuscript.tex
second round revisions at ecography

\paragraph*{FAIR data}
The FAIR principles represent guidelines for making data more persistent, findable, and well-documented. Structuring the data as an R package ensures that metadata and data are packaged together, where R manual files contain column names and data descriptions for each data product (\textit{Findable}). All code to take data from the raw data (\texttt{data-raw} folder) to the end product .RData and .csv files is contained in the versioned R data package, and integration with Zenodo (\texttt{https://doi.org/10.5281/zenodo.5821896}) provides a DOI for each release (\textit{Accessible}). Apart from providing data in these multiple formats, user access is aided by structuring the data as a package in a very popular computing language among biologists (and other folks too) and providing all code for data processing and serving in an open and public-facing repository (\textit{Interoperable}). Having all code and data in a streamlined, open, and versioned format, serving the data through an interactive web portal, and publishing this software note collectively serve to promote the use of this data resource (\textit{Reusable}).

@@

Differences in features across the data on different pathogen types (e.g., \texttt{?nematodes} relative to \texttt{?viruses}) make combining these data non-straightforward, without a degree of loss of information. We provide some example code in the package \texttt{vignette} on how to go about combining or linking the data across types, with the caveats of information loss, and have standardized some key column names across the different data products. Further, we have documented each data resource using $R$ package documentation, allowing the metadata of each data product to be examined directly from R using the \texttt{help()} function or the question mark notation (e.g., \texttt{?viruses}).

>>>>

\paragraph*{FAIR data}
The FAIR principles represent guidelines for making data more persistent, findable, and well-documented. Structuring the data as an R package ensures that metadata and data are packaged together, where R manual files contain column names and data descriptions for each data product (\textit{Findable}). All code to take data from the raw data (\texttt{data-raw} folder) to the end product .RData and .csv files is contained in the versioned R data package, and integration with Zenodo (\texttt{https://doi.org/10.5281/zenodo.5821896}) provides a DOI for each release (\textit{Accessible}). Metadata are available in redundant forms, both from within the R package as \texttt{man} files, and in the project README file such that installation of the package (or navigation into the \texttt{man} folder) is not necessary. Apart from providing data in these multiple formats, user access is aided by structuring the data as a package in a very popular computing language among biologists (and other folks too) and providing all code for data processing and serving in an open and public-facing repository (\textit{Interoperable}). Having all code and data in a streamlined, open, and versioned format, serving the data through an interactive web portal, and publishing this software note collectively serve to promote the use of this data resource (\textit{Reusable}).

@@

Differences in features across the data on different pathogen types (e.g., \texttt{?nematodes} relative to \texttt{?viruses}) make combining these data non-straightforward, without a degree of loss of information. We provide some example code in the package \texttt{vignette} on how to go about combining or linking the data across types, with the caveats of information loss, and have standardized some key column names across the different data products. Further, we have documented each data resource using $R$ package documentation, allowing the metadata of each data product to be examined directly from R using the \texttt{help()} function or the question mark notation (e.g., \texttt{?viruses}). These same metadata are also provided in the README file in the top-level of the GitHub repository.
